Single Bivalent Level Analysis

### 2020-01-23 doc/Bivalent_Analysis.Rmd
I think I got the code for isolating long biv data set debugged. v tedious

#then this DF has to be summarized into categories
chrm.class.category <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                             mean.percent.type = mean(value)
                             #sum is the wrong!
                             #used average percentage, might consider a different summary
)

@@

chrm.class.barplot <- ggplot(data = chrm.class.category, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
chrm.class.barplot

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
show.the.cells_females_musc

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_Musc_f

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
show.the.cells_male_musc

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_Musc_m

#the min is pretty stable

@@

@@

for(  i in 1:length( long_BivData_set99$fileName)  ){
  long_BivData_set99$cell.Q4[i] <- Biv_Data_mean_table[ which(Biv_Data_mean_table$fileName == long_BivData_set99$fileName[i]), ]
}

for(  k in 1:length( long_BivData_set99$fileName)  ){
long_BivData_set99$above.Q4[k] <- ifelse( (long_BivData_set99$cell.Q4 < long_BivData_set99$chromosomeLength), "yes", "no")
}
#This isn't working either

#for(i in 1:length( long_BivData_set$fileName)  ){ 
  # print(i)
#  current_obj_fileName <- long_BivData_set$fileName[i]
#  current_obj_SC_length <- long_BivData_set$chromosomeLength[i]  
#  current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == current_obj_fileName) ]
# current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == long_BivData_set$fileName[i]) ]
#  long_BivData_set$above.Q4 <-ifelse( (current.Q4 < current_obj_SC_length), 1, 0)
#  print( paste( c(current.Q4, current_obj_fileName, current_obj_SC_length), sep = "_" ) )
#}
#this loop isn't working

@@

@@

2) normalized CO positions (sedell and Kirkpatrick).

@@

@@

Sex is a significant effect for SC length
The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)

@@

There are sex effects for the lme() models for both the ABS and PER IFD measures. There isn't evidence of Random effects from the lmer() models.

@@

> Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?

@@

> Is there evidence for the alternative IFD meansure

INSERT THE SISCOTEN ESTIMATES!!

```{r siscoten.test, echo=FALSE}

>>>>

#then this DF has to be summarized into categories
chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                             mean.percent.type = mean(value)
                             #sum is the wrong!
                             #used average percentage, might consider a different summary
)

@@

chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
chrm.class.barplot

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")

@@

#adding the extra cols, messed up this part

@@

long_BivData_set99 <- long_BivData_set99[1:50,]
#3 unique fileNames, in abbridged DF

ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))

newfile = strsplit(i, "_")[[1]][1:7]

Biv_Data_mean_table[grepl(newfile, Biv_Data_mean_table$fileName )]

#i is obj.id, this returns a table of all falses -- because it's look for the whole string

o<- strsplit(i, "_")

Biv_Data_mean_table$fileName[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
#above returns the BivData, filename which the grepl matches

Biv_Data_mean_table$Qar.4[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
#this returns the Q4

#this will put the cell's Q4
#I think I need a counter
counter = 1

length(duplicated(long_BivData_set99$Obj.ID))#there are 2605 duplicated Obj.ID
anyDuplicated(long_BivData_set99$Obj.ID)
dup.curated <- long_BivData_set99[duplicated(long_BivData_set99$Obj.ID),]#65 duplicated obj.ID
long_BivData_set99 <- long_BivData_set99 %>% distinct()

#REDO
#61 unique biv.id

#this must be charecter for str split

long_BivData_set99$Obj.ID <- as.character(long_BivData_set99$Obj.ID )
long_BivData_set99.mini <- long_BivData_set99[905:2605,]

long_BivData_set99.mini$Obj.ID <- paste(long_BivData_set99.mini$fileName, long_BivData_set99.mini$boxNumber, sep = "_")

anyDuplicated(long_BivData_set99.mini$Obj.ID)#1093
dup.mini.curated <- long_BivData_set99.mini[duplicated(long_BivData_set99.mini$Obj.ID),]

#do a pipe remove
long_BivData_set99.mini <- long_BivData_set99.mini[ ! long_BivData_set99.mini$fileName %in% dup.mini.curated$fileName, ]
#think this removes the duplicated rows

#IT COULD BE THAT THERE are filenames in the Biv_Data_mean_table which don't have the

#make sure the biv.mini only have filenames which are also in the mean teabl
long_BivData_set99.mini <- long_BivData_set99.mini[ ! long_BivData_set99.mini$fileName %in% dup.mini.curated$fileName, ]

long.data.set.cells <- al.15_bivs

long_BivData_set99.mini <- long_BivData_set99.mini[long_BivData_set99.mini$fileName %in% long.data.set.cells$fileName, ]

long_BivData_set99.mini <- long_BivData_set99.mini[long_BivData_set99.mini$fileName %in% Biv_Data_mean_table$fileName, ]
#removed fileNames which don't appear in Biv_Data_mean_table (which has the Q4)

#long_BivData_set99.mini <- long_BivData_set99[905:2605,]

for(  i in 1:length( unique(long_BivData_set99.mini$Obj.ID) ) ){
  newfile.str = strsplit( long_BivData_set99.mini$Obj.ID[i], "_")[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
  #the above can produe mulitple Q4

  long_BivData_set99.mini$cell.Q4.99[i] <- the.Q4
  #these are just using the last index,

  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]

  long_BivData_set99.mini$Q4.FN[i] <- the.FN

  print(c(i, the.FN, the.Q4))

}
#some kind of error at 903 -- the loop should continue past it

long_BivData_set99.mini$FN2 <- long_BivData_set99.mini$fileName
#mini, looks good so far

long_BivData_set99.mini$long.Biv <- ifelse( (long_BivData_set99.mini$chromosomeLength >= long_BivData_set99.mini$cell.Q4), "yes", "no" )

long_BivData_set99.mini$wtf <- long_BivData_set99.mini$chromosomeLength - long_BivData_set99.mini$cell.Q4

real.long.bivData <- long_BivData_set99.mini[long_BivData_set99.mini$long.Biv == "yes",]
table(real.long.bivData$fileName)

#!!! omg it finally worked!~!!
#OMFG

#OG try
#for(  i in unique(long_BivData_set99$Obj.ID) ){
#  newfile = strsplit(i, "_")[[1]][1:7]
#  the.Q4 <- Biv_Data_mean_table$Qar.4[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
#  long_BivData_set99$cell.Q4.99[counter] <- the.Q4
#  #these are just using the last index, 
#  the.FN <- Biv_Data_mean_table$fileName[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
#  long_BivData_set99$Q4.FN[counter] <- the.FN
#  print(c(i, the.FN, the.Q4))
#  counter = counter+1
#}
#not sure what the final error is in regards to
#the file name matches are wronge

@@

Ugggg, still something wrong,
9apr18_1feb18_LEW_f1_sp1_9_rev

 0%    25%    50%    75%   100% 
 77.00 114.50 136.00 152.75 212.00

Started compiling the long_biv data set, some cells have more than I expect:

13apr15_1apr15_PWD_f2_sp1_4_rev, 5aug17_14jul17_MSM_f1_sp1_12_rev, 5aug17_14jul17_MSM_f1_sp1_26_rev, 30jun17_4jan17_LEW_f3_sp1_1_rev, 8feb17_4jan17_LEW_f1_sp1_5_rev, 8feb17_4jan17_LEW_f1_sp1_3_rev, 9apr18_1feb18_LEW_f1_sp1_9_rev, 8jun15_23apr15_PWD_f2_sp1_17_rev

They have like ~10, it might have to do with not haveing = in the >=

Qar.4 = quantile(chromosomeLength)[[4]],
quantile(Curated_BivData$chromosomeLength[Curated_BivData$fileName == "8feb17_4jan17_LEW_f1_sp1_3_rev"]) #148.00
Curated_BivData$chromosomeLength[Curated_BivData$fileName == "8feb17_4jan17_LEW_f1_sp1_3_rev"]

9apr18_1feb18_LEW_f1_sp1_9_rev

the correct Q4 is 148, but the value in the Q4 col is 116, and the filName listed is completely wrong

@@

2.A) normalized CO positions (sedell and Kirkpatrick).

  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.

@@

D) Sis-co-ten predictions fall into

@@

- Sex is a significant effect for SC length
The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)

- According to anova, sex effect explains most of the variance in single bivalent SC lengths.

@@

There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.

@@

> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?

@@

> Prediction B, Is there evidence for the alternative IFD meansure
